---
title: "evs-in-md.rmd"
author: "Rina"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load Libraries
```{r}

library(httr)
library(jsonlite)
library(tidyverse)
library(leaflet)
library(shiny)
library(sf)
library(tidygeocoder)
library(tidycensus)
library(ggplot2)


```
##Accessing the U.S. Department of Energy's Electric Vehicle Station API

#We saved the API Key in our R environment first.

```{r}
#We construct the URL we need to get into the API. The Sys.getenv references the API key that I

my_url <- paste0("https://developer.nrel.gov/api/alt-fuel-stations/v1.json?&api_key=", 
                 Sys.getenv("CHARGING_API_KEY"),
                 "&fuel_type=ELEC&state=MD")

res = GET(my_url)
data = fromJSON(rawToChar(res$content))
names(data)
stations <- data$fuel_stations
```

#Check to see if results total matches the number of rows I got in stations dataframe.
```{r}

results <- data$total_results

```

#Select columns of interest.
```{r}

station_selects <- stations %>%
  select(access_code, access_days_time, date_last_confirmed, fuel_type_code, id, open_date, restricted_access, station_name, facility_type, latitude, longitude, city, state, street_address, zip, ev_connector_types)

#Some light cleaning
clean_stations <- station_selects%>%
  mutate(latitude=ifelse(id==195872,39.3396817,latitude),
         longitude=ifelse(id==195872,-76.4812393,longitude),
         latitude=ifelse(id==193781,39.0489226, latitude),
         longitude=ifelse(id==193781,-76.9579801, longitude),
         latitude=ifelse(id==250383,38.9898679, latitude),
         longitude=ifelse(id==250383,-77.096218, longitude))

  
```
#Note: This takes a while! The data in the app was extracted on May 13.

#Get county data from from Electric Vehicle API
```{r}

geo_stations <- clean_stations %>%
  reverse_geocode(lat = latitude, long = longitude, method = 'osm', full_results = TRUE)


geo_stations <- geo_stations%>%
  select(access_code, access_days_time, date_last_confirmed, fuel_type_code, id, open_date, restricted_access, station_name, facility_type, latitude, longitude, city, state, street_address, zip, county)%>%
  rename("state" = "state...13")

geo_stations$county <- geo_stations$county %>% replace_na('Baltimore City')

write.csv(reverse_all, file='charging513.csv', row.names=FALSE)

```


#Grabbing Census Layers Examples
#You'll need a Census API Key to get this running.
```{r}

#Income Tract
md_income_tract <-get_acs(geography = "tract", 
                             variables = "B19013_001", 
                             state = "MD", 
                             geometry = TRUE)%>%
                      rename(income = estimate)

#Create Shapefile
st_write(md_income_tract, "incomes.shp")

#County Populations
pop_county <- get_acs(geography = "county", 
                              variables = "B01003_001", 
                              state = "MD",
                              geometry = TRUE)
#Create Shapefile
st_write(pop_county, "pop_county.shp")

```




